### 2012-10-01 MSteenhagen-CV.tex
New CV created. Needs polishing.

%%% A template to produce a nice-looking Curriculum Vitae.
%%% Kieran Healy <kjhealy@gmail.com>
%%% Most recent version is at http://kjhealy.github.com/kjh-vita
%%%
%%% ------------------------------------------------------------------------
%%% Requirements (should be included in a modern tex distribution):
%%% ------------------------------------------------------------------------
%%% xelatex
%%% fontspec.sty
%%% hyperrref.sty
%%% xunicode.sty
%%% color.sty
%%% url.sty
%%% fancyhdr.sty
%%%
%%% ------------------------------------------------------------------------
%%% Optional
%%% ------------------------------------------------------------------------
%%% git
%%% vc.sty
%%% revnum.sty
%%% Fonts
%%%
%%% ------------------------------------------------------------------------
%%% Note
%%%------------------------------------------------------------------------
%%% Because this is a hand-tweaked file, be on the look out for \medksip, 
%%% \bigskip and \newpage commands here and there, which are used to balance
%%% the layout or avoid widows & orphans, etc. You should of course add or 
%%% remove these as needed.
%%%------------------------------------------------------------------------

\documentclass[11pt]{article}

%%%------------------------------------------------------------------------
%%% Metadata
%%%------------------------------------------------------------------------

%% Change as needed. Or just add me as a coauthor. Only some of these are 
%% used below in the hyperref declaration and address banner section.
\def\myauthor{Maarten Steenhagen}
\def\mytitle{Vita}
\def\mycopyright{\myauthor}
\def\mykeywords{}
\def\mybibliostyle{plain}
\def\mybibliocommand{}
\def\mysubtitle{}
\def\myaffiliation{University College London}
\def\myaddress{Department of Philosophy}
\def\myemail{m.steenhagen.09@ucl.ac.uk}
\def\myweb{http://ucl.ac.uk/\string~uctymst}
\def\myblog{http://maarten.steenhagen.nl}
\def\myphone{+44 (0)777 1193717}
\def\myfax{~}
\def\myversion{}
\def\myrevision{}

\def\myaffiliation{University College London}
\def\myauthor{Maarten Steenhagen}
\date{} % not used (revision control instead)
\def\mykeywords{Maarten, Steenhagen, Maarten Steenhagen, Vita, CV, Resume, Philosophy}

%%%------------------------------------------------------------------------  
%%% Git version tracking 
%%%------------------------------------------------------------------------

%% If you don't use git or the vc package (from CTAN), comment this out.
%% If you comment it out, be sure to remove the \rfoot comment below, too.
%\input{vc}

%%%------------------------------------------------------------------------
%%% Required style files
%%%------------------------------------------------------------------------
\usepackage{url,fancyhdr}
%%\usepackage{revnum} % for reverse-numbered publications (revnumerate environment) if needed.

%% needed for xelatex to work
\usepackage{fontspec}
\usepackage{xunicode}

%% color for the links 
\usepackage{color}

%% hyperlinks
\usepackage[xetex, 
	colorlinks=false,
	plainpages=false,
  	pdfpagelabels,
  	bookmarksnumbered,
  	pdftitle={\mytitle},
  	pagebackref,
  	pdfauthor={\myauthor},
  	pdfkeywords={\mykeywords}
  	]{hyperref}

%%%------------------------------------------------------------------------
%%% Document
%%%------------------------------------------------------------------------
\begin{document}

%% Choose fonts for use with xelatex
%% Minion and Myriad are widely available, from Adobe. 
%% Pragmata is available to buy at http://www.fsd.it/fonts/pragma.htm
%% and is worth every penny. Any good monospace font will work fine, though.
%% Consolas or inconsolata are good alternatives.
\setromanfont[Mapping={tex-text},Numbers={OldStyle},Ligatures={Common}]{Brill} 
\newfontfamily\margin[Mapping=tex-text,Colour=AA0000, Letters=SmallCaps]{Futura}
\setmonofont[Mapping=tex-text,Scale=0.9]{Helvetica}

% \newcommand{\amps}{\amper{\&}}

%%%------------------------------------------------------------------------
%%% Local commands
%%%------------------------------------------------------------------------

%% Marginal header
%% Note: as the document goes on you may need to introduce a (gradually increasing)
%% \vspace element to keep the marginal header pleasingly aligned with the first 
%% item in the body text. Like this: \marginhead{{\vskip 0.4em}Grants}, or 
%% \marginhead{{\vskip 0.8em}Service}. Experiment as needed.
\newcommand{\marginhead}[1]{\marginpar{\margin{{\footnotesize\vspace{-1em}\flushright #1}}}}

%% custom ampersand (font consistent with the one chosen above)
\newcommand{\amper}{{\fontspec[Scale=.85]{Baskerville Italic}\selectfont\&\,}}

%% No bullets on labels
\renewcommand{\labelitemi}{~}

%% Custom hanging indent for vita items
\def\ind{\hangindent=1 true cm\hangafter=1 \noindent}
%\def\ind{\hangindent=18pt\hangafter=1 \noindent}
\def\labelitemi{~}
\renewcommand{\labelitemii}{~}

%%%------------------------------------------------------------------------
%%% Page layout
%%%------------------------------------------------------------------------
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\fancyhead{}
\fancyfoot{}
\rhead{{\scriptsize\thepage}}

%% git revision control footer 
%\rfoot{\texttt{\scriptsize \VCRevision\ on \VCDateTEX}} % git revision info inserted via external script -- see docs for vc package for details. comment out this line if you're not using vc, and also remove the \input{vc} line above.

%%%------------------------------------------------------------------------
%%% Address and contact block
%%%------------------------------------------------------------------------
\begin{minipage}[t]{2.95in}
 \flushright {\footnotesize \href{http://ucl.ac.uk/philosophy}{Department of Philosophy} \\ University College London \\ Gower Street \\ \vspace{-0.05in} London, \textsc{wc1e 6bt}}

\end{minipage}
\hfill     
%\begin{minipage}[t]{0.0in}
% dummy (needed here)
%\end{minipage}
\hfill
\begin{minipage}[t]{1.7in}
  \flushright \footnotesize Phone: \myphone \\  
  {\scriptsize  \texttt{\href{mailto:\myemail}{\myemail}}} \\
  {\scriptsize  \texttt{\href{\myweb}{\myweb}}}
	{\scriptsize  \texttt{\href{\myblog}{\myblog}}}
\end{minipage}

\vfil

%% Name 
\noindent{\Large {\textsc{maarten steenhagen}}}
\reversemarginpar

\bigskip

%% Appointments
\medskip

\marginhead{Current Position}

\noindent\emph{University College London \vspace{0.01in}}

\ind Doctoral Candidate (PhD) in Philosophy.

\ind Dissertation research on the nature of human experience of pictorial representations.

\bigskip

\marginhead{AoS}

\ind Philosophy of Psychology; Philosophy of Mind \amper Language; Aesthetics; Metaethics.

\bigskip
\bigskip

\marginhead{AoC}

\ind Metaphysics; History of Twentieth Century Philosophy; Political Philosophy.

\bigskip
\bigskip
\vfil 
\marginhead{Education \newline (Past)}

\noindent\emph{University College London \vspace{0.01in}}

% \ind 2012--Present. Doctoral Candidate (PhD).

\ind 2012. MPhil Stud. research student. Dissertation: \emph{The Collective Image}. \vspace{-0.1in}

\bigskip

\noindent\emph{University of Utrecht\vspace{0.02in}}

\ind 2009. B.A., Philosophy.

\medskip

\noindent\emph{ARTeZ School of Fine Art\vspace{0.02in}}

\ind 2005. BFA., Fine Art.

\bigskip

\marginhead{Appointments}

\noindent\emph{Heythrop College London \vspace{0.01in}}

\ind 2009--Present. Tutorial Assistant in Philosophy and Psychology.

\ind 2009--Present. Research Assistant in Philosophy.

\medskip

\noindent\emph{University of Kent \vspace{0.01in}}

\ind 2011--2012. Assistant Lecturer, History and Philosophy of Art.

\medskip

\noindent\emph{University College London \vspace{0.01in}}

\ind 2011-2012. Teaching Assistant in Philosophy.

\medskip

\noindent\emph{Utrecht School of Arts \vspace{0.01in}}

\ind 2008-2009. Assistant Lecturer in Art Theory and Cultural History.

\medskip

\noindent\emph{University of Utrecht \vspace{0.01in}}

\ind 2007-2008. Teaching Assistant in Aesthetics and Moral Psychology.

\bigskip

\newpage

%% Presentations
\marginhead{{\vskip 0.4em}Talks}
\medskip

\ind 2012. ``How the five die: Response to Lamb.'' University College London, 2012 Graduate Conference in Philosophy. (Response paper.)

\ind 2012. ```Non-conceptualised depiction.'' The Queen's College Oxford, Annual conference of the British Society of Aesthetics. September.

\ind 2012. ``Researching the Self.'' St. John's College Oxford, \emph{Art as a Mode of Enquiry}.

\ind 2011. ``Seeing what another would.'' Utrecht University, \emph{Expert Meeting on Philosophical Aesthetics}.

\ind 2011. ``Representing the Trinity.'' University of Ghent, Annual conference of the Dutch Society of Aesthetics.

\ind 2011. ``Truly political? A psychoanalytic understanding of engaged art.'', University of Leeds, \emph{Working Through Psychoanalysis}.

\ind 2010. ``Photography and dull causation: Response to Atencia-Linares.'' University College London, 2010 Graduate Conference in Philosophy. (Response paper.)

\ind 2010. ``Explaining the Ugly: Disharmony and Unrestrained Cognition in Kant.'' Institute of Philosophy, London-Berkeley Graduate Conference in Philosophy.

%\end{revnumerate}
\bigskip

\marginhead{{\vskip 0.4em}Grants and \newline Awards}
\medskip

\ind 2012. British Society of Aesthetics PhD Studentship Award. (£36.000)

\ind 2012. Prins Bernhard Cultuurfonds (€10.000). \emph{Not used.}

\ind 2012. Hendrik Müller Vaderlandsch Fonds (€5.000). \emph{Not used.}

\ind 2010. Dutch Association for Aesthetics Essay Contest, 2$^{\textrm{\tiny{nd}}}$ Prize.

\ind 2009. Huygens Scholarship Programme, Nuffic, The Hague (€30.000).

\bigskip

\marginhead{{\vskip 0.9em}Service to the \newline Profession}
\medskip

\medskip

\ind 2012--Present. Editor of the \emph{Postgraduate Journal of Aesthetics.}

\ind 2010--Present. Co-Organiser London Aesthetics Forum, Institute of Philosophy, London.

\ind 2012--Present. Website Editor London Philosophy and Psychoanalysis Group.

\ind 2012. Assistant Organiser European Society for Philosophy and Psychology (ESPP) Annual Conference 2012, Institute of Philosophy, London.

\ind 2011. UCL Representative for the British Postgraduate Philosophy Association.

\ind 2007-2009. Editor of \emph{Cimedart}, Philosophy Journal of the University of Amsterdam.

% %% Publications
% \marginhead{{\vskip 0.3em}Publications}
% \medskip
% \noindent\emph{Book \vspace{0.01in}}
% 
% \ind  Kieran Healy. 2006. \emph{\href{http://www.lastbestgifts.com}{Last Best Gifts: Altruism and the Market for Human Blood and Organs}}. Chicago:~University of Chicago Press. \vspace{-0.075in}
%  
% \normalsize
% 
% \bigskip
% \noindent\emph{Journal articles \vspace{0.05in}}
%  
% %% Use revnumerate environment if numbered publications are needed. 
% %% (Include it above in the preamble).
% %% \renewcommand{\labelenumi}{\textsc{a}\theenumi.}
% %% \begin{revnumerate}
% 
% \ind Marion Fourcade and Kieran Healy. 2007. ``\href{http://arjournals.annualreviews.org/doi/pdf/10.1146/annurev.soc.33.040406.131642}{Moral Views of Market Society}.'' \emph{Annual Review of Sociology} 33:~285--311. 
% 
% \ind Kieran Healy. 2006. ``\href{http://www.kieranhealy.org/files/papers/presumed-consent.pdf}{Do Presumed Consent Laws Raise Organ Procurement Rates?}'' \emph{DePaul Law Review} 55:~1017--1043. 
% 
% \ind Kieran Healy. 2004. ``\href{http://www.kieranhealy.org/files/papers/asr2004.pdf}{Altruism as an Organizational Problem: The Case of Organ Procurement}.''
%  \emph{American Sociological Review} 69:~387--404.
% 
% \ind Kieran Healy. 2002. ``\href{http://www.kieranhealy.org/files/papers/jpp.pdf}{Digital Technology and Cultural Goods}.'' \emph{Journal of
%   Political Philosophy} 10:~478--500. 
% 
% \ind Kieran Healy. 2002. ``\href{http://www.kieranhealy.org/files/papers/culture-newecon.pdf}{What's New for Culture in the New Economy?}'' \emph{Journal
%   of Arts Management, Law and Society} 32:~86--103. 
% 
% 
% \ind Kieran Healy. 2000. ``\href{http://www.kieranhealy.org/files/papers/embed-alt.pdf}{Embedded Altruism: Blood Collection Regimes and the European
%   Union's Donor Population}.'' \emph{American Journal of Sociology} 105:~1633--\,1657.
% 
% 
% \ind Kieran Healy. 1999.  ``\href{http://www.kieranhealy.org/files/papers/ts.pdf}{The Emergence of \textsc{hiv} in the U.S. Blood Supply: Organizations,
%   Obligations and the Management of Uncertainty}.'' \emph{Theory and Society} 28:~529--558.
% 
% 
% \ind Bruce Western and Kieran Healy. 1999. ``\href{http://www.kieranhealy.org/files/papers/western-healy99.pdf}{Explaining the OECD Wage Slowdown:
% Recession or Labor Decline?}''  \emph{European Sociological Review}
% 15:~233--249.
% 
% 
% \ind  Kieran Healy. 1998.  ``\href{http://www.kieranhealy.org/files/papers/constraint98.pdf}{Conceptualising Constraint: Mouzelis, Archer and the concept of
%   social structure}.'' \emph{Sociology} 32:~509--522. 
% 
% %\end{revnumerate}
% %\newpage
% \bigskip
% 
% \noindent\emph{Book chapters \vspace{0.05in}}
% % \renewcommand{\labelenumi}{\textsc{c}\theenumi.}
% % \begin{revnumerate}
% 
% \ind Kieran Healy. Forthcoming. ``Social Structure, Gifts and Norms in \emph{The Story of Qiu Ju},'' in \emph{Norms and Values: The Role of Social Norms as Instruments of Value Realisation}, edited by Michael Baurmann, Geoffrey Brennan, Robert E. Goodin and Nicholas Southwood. Baden Baden: Nomos Verlag.
% 
% \ind Kieran Healy. 2008. ``Lo Scambio di Doni,'' in \emph{L'Altro in Me: Dono del Sangue fra Culture, Pratiche e Identità}, edited by Annamaria Fantauzzi. Milano: \textsc{avis} Nazionale, 36--42. 
% 
% \ind Kieran Healy. 2007. ``Sociology,'' in \emph{A Companion to Contemporary Political Philosophy} (Second Edition), edited by Robert E. Goodin, Philip Pettit and Thomas Pogge. New York: Blackwell, 90--122. 
% 
% \ind  Kieran Healy. 2004. ``Sacred Markets and Secular Ritual in the Organ Transplant
% Industry,'' in \emph{The Sociology of the Economy}, edited by Frank Dobbin. New
% York: Russell Sage, 336--359.
% 
% \ind Bruce Western and Kieran Healy.  2001.  ``Wage Growth and Labor Decline in the
% Industrialized Democracies, 1965--1993,'' in \emph{Unemployment in the New
%   Europe}, edited by Nancy Bermeo. New York: Cambridge University Press,
% 121--144.
% 
%  \ind  Kieran Healy. 1998. ``The New Institutionalism and Irish Social Policy,'' in
% \emph{Social Policy in Ireland: Principles, Practice and Problems}, edited by
% Se\'{a}n Healy and Brigid Reynolds. Dublin: Oak Tree Press, 59--83.
% 
% %\end{revnumerate}
% 
% \bigskip 
%  
% %\newpage
% \noindent\emph{Reviews and other occasional writing \vspace{0.05in}}
% 
% %\renewcommand{\labelenumi}{\textsc{r}\theenumi.}
% %\begin{revnumerate}
% \ind Kieran Healy. 2009. \href{http://www.journals.uchicago.edu/doi/pdf/10.1086/605757}{Review} of \emph{Pop Finance} by Brooke Harrington. \emph{American Journal of Sociology} 115:~309--12.
% 
% \ind Kieran Healy. 2007. ``\href{http://www.kieranhealy.org/files/reviews/socforum-article.pdf}{A Wealth of Notions}.'' Review
% Essay on \emph{Freakonomics} by Steven Levitt and Stephen Dubner. \emph{Sociological Forum} 22:~119--125. 
% 
% \ind  Kieran Healy. 2006. ``Sacred.'' In \emph{International Encyclopedia of Economic Sociology}, edited by Jens Beckert and Milan Zafirovski. London: Routledge, 588--589. 
% 
% \ind  Kieran Healy. 2005. \href{http://www.kieranhealy.org/files/reviews/klinenberg.pdf}{Review} of \emph{Heat Wave} by Eric Klinenberg. \emph{Imprints} 8:~283--289.
% 
% \ind  Kieran Healy. 2005. \href{http://www.kieranhealy.org/files/reviews/fox.pdf}{Review} of \emph{An R and S-PLUS Companion to Applied Regression}, by John Fox. \emph{Sociological Methods and Research} 34:~137--140.
% 
% \ind  Kieran Healy. 2005. ``\href{http://www.kieranhealy.org/files/reviews/henwood.pdf}{After the Ball was Over}.'' Review of \emph{After the New
%   Economy} by Doug Henwood. \emph{European Economic Sociology Newsletter} 6 (3):~37--38.
% 
% \ind  Kieran Healy. 2004. ``The Abundance of the Heart.'' Review of \emph{A Man After His Own Heart} by Charles
% Siebert. \emph{Australian Review of Public Affairs}, October. Reprinted in the
% \emph{Australian Financial Review}, October 22nd.
% 
% \ind  Kieran Healy. 2003. \href{http://links.jstor.org/sici?sici=0094-3061%28200303%2932%3A2%3C252%3AAEATOC%3E2.0.CO%3B2-F}{Review} of \emph{Against Essentialism} by Stephan Fuchs.
% \emph{Contemporary Sociology} 32:~252--254.
% 
% \ind  Kieran Healy. 2001. Review of \emph{Market Society} by Don Slater and Fran Tonkiss. \emph{European
%       Economic Sociology Newsletter} 3 (1):~31--32.
% 
% 
% \ind  Kieran Healy. 1999. \href{http://links.jstor.org/sici?sici=0094-3061%28199903%2928%3A2%3C210%3ACPRCAS%3E2.0.CO%3B2-C}{Review} of \emph{Comparative Politics} by Mark Irving Lichbach and Alan S.
% Zuckerman. \emph{Contemporary Sociology} 28:~210--211.
% 
% 
% \ind  Kieran Healy. 1998. \href{http://www.kieranhealy.org/files/reviews/rule.pdf}{Review} of \emph{Theory and Progress in Social Science} by James Rule.  \emph{Theory
%   and Society}, 27:~435--441.
% 
% % %\end{revnumerate}
%  \bigskip

\end{document}