### 2012-10-02 MSteenhagen-CV.tex
Removed foreign entries.

%%% A template to produce a nice-looking Curriculum Vitae.
%%% Orginally produced by Kieran Healy <kjhealy@gmail.com>
%%% This version is adapted (only minimally) by Maarten Steenhagen.
%%% 
%%% ------------------------------------------------------------------------
%%% Requirements (should be included in a modern tex distribution):
%%% ------------------------------------------------------------------------
%%% xelatex
%%% fontspec.sty
%%% hyperrref.sty
%%% xunicode.sty
%%% color.sty
%%% url.sty
%%% fancyhdr.sty
%%%
%%% ------------------------------------------------------------------------
%%% Optional
%%% ------------------------------------------------------------------------
%%% git
%%% vc.sty
%%% revnum.sty
%%% Fonts
%%%
%%% ------------------------------------------------------------------------
%%% Note
%%%------------------------------------------------------------------------
%%% Because this is a hand-tweaked file, be on the look out for \medksip, 
%%% \bigskip and \newpage commands here and there, which are used to balance
%%% the layout or avoid widows & orphans, etc. You should of course add or 
%%% remove these as needed.
%%%------------------------------------------------------------------------

\documentclass[11pt]{article}

%%%------------------------------------------------------------------------
%%% Metadata
%%%------------------------------------------------------------------------

%% Change as needed. Or just add me as a coauthor. Only some of these are 
%% used below in the hyperref declaration and address banner section.
\def\myauthor{Maarten Steenhagen}
\def\mytitle{Vita}
\def\mycopyright{\myauthor}
\def\mykeywords{}
\def\mybibliostyle{plain}
\def\mybibliocommand{}
\def\mysubtitle{}
\def\myaffiliation{University College London}
\def\myaddress{Department of Philosophy}
\def\myemail{m.steenhagen.09@ucl.ac.uk}
\def\myweb{http://ucl.ac.uk/\string~uctymst}
\def\myblog{http://maarten.steenhagen.nl}
\def\myphone{+44 (0)777 1193717}
\def\myfax{~}
\def\myversion{}
\def\myrevision{}

\def\myaffiliation{University College London}
\def\myauthor{Maarten Steenhagen}
\date{} % not used (revision control instead)
\def\mykeywords{Maarten, Steenhagen, Maarten Steenhagen, Vita, CV, Resume, Philosophy}

%%%------------------------------------------------------------------------  
%%% Git version tracking 
%%%------------------------------------------------------------------------

%% If you don't use git or the vc package (from CTAN), comment this out.
%% If you comment it out, be sure to remove the \rfoot comment below, too.
%\input{vc}

%%%------------------------------------------------------------------------
%%% Required style files
%%%------------------------------------------------------------------------
\usepackage{url,fancyhdr}
%%\usepackage{revnum} % for reverse-numbered publications (revnumerate environment) if needed.

%% needed for xelatex to work
\usepackage{fontspec}
\usepackage{xunicode}

%% color for the links 
\usepackage{color}
\definecolor{grey}{cmyk}{0,1,0,0.8}
%% hyperlinks
\usepackage[xetex, 
	colorlinks=true,
	urlcolor=grey,
	plainpages=false,
  	pdfpagelabels,
  	bookmarksnumbered,
  	pdftitle={\mytitle},
  	pagebackref,
  	pdfauthor={\myauthor},
  	pdfkeywords={\mykeywords}
  	]{hyperref}

%%%------------------------------------------------------------------------
%%% Document
%%%------------------------------------------------------------------------
\begin{document}

%% Choose fonts for use with xelatex
%% Minion and Myriad are widely available, from Adobe. 
%% Pragmata is available to buy at http://www.fsd.it/fonts/pragma.htm
%% and is worth every penny. Any good monospace font will work fine, though.
%% Consolas or inconsolata are good alternatives.
\setromanfont[Mapping={tex-text},Numbers={OldStyle},Ligatures={Common}]{Brill} 
\newfontfamily\margin[Mapping=tex-text,Colour=AA0000, Letters=SmallCaps]{Futura}
\setmonofont[Mapping=tex-text,Scale=0.9]{Helvetica}

% \newcommand{\amps}{\amper{\&}}

%%%------------------------------------------------------------------------
%%% Local commands
%%%------------------------------------------------------------------------

%% Marginal header
%% Note: as the document goes on you may need to introduce a (gradually increasing)
%% \vspace element to keep the marginal header pleasingly aligned with the first 
%% item in the body text. Like this: \marginhead{{\vskip 0.4em}Grants}, or 
%% \marginhead{{\vskip 0.8em}Service}. Experiment as needed.
\newcommand{\marginhead}[1]{\marginpar{\margin{{\footnotesize\vspace{-1em}\flushright #1}}}}

%% custom ampersand (font consistent with the one chosen above)
\newcommand{\amper}{{\fontspec[Scale=.85]{Baskerville Italic}\selectfont\&\,}}

%% No bullets on labels
\renewcommand{\labelitemi}{~}

%% Custom hanging indent for vita items
\def\ind{\hangindent=1 true cm\hangafter=1 \noindent}
%\def\ind{\hangindent=18pt\hangafter=1 \noindent}
\def\labelitemi{~}
\renewcommand{\labelitemii}{~}

%%%------------------------------------------------------------------------
%%% Page layout
%%%------------------------------------------------------------------------
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\fancyhead{}
\fancyfoot{}
\rhead{{\scriptsize\thepage}}

%% git revision control footer 
%\rfoot{\texttt{\scriptsize \VCRevision\ on \VCDateTEX}} % git revision info inserted via external script -- see docs for vc package for details. comment out this line if you're not using vc, and also remove the \input{vc} line above.

%%%------------------------------------------------------------------------
%%% Address and contact block
%%%------------------------------------------------------------------------
\begin{minipage}[t]{2.95in}
 \flushright {\footnotesize \href{http://ucl.ac.uk/philosophy}{Department of Philosophy} \\ University College London \\ Gower Street \\ \vspace{-0.05in} London, \textsc{wc1e 6bt}}

\end{minipage}
\hfill     
%\begin{minipage}[t]{0.0in}
% dummy (needed here)
%\end{minipage}
\hfill
\begin{minipage}[t]{1.7in}
  \flushright \footnotesize Phone: \myphone \\  
  {\scriptsize  \texttt{\href{mailto:\myemail}{\myemail}}} \\
  {\scriptsize  \texttt{\href{\myweb}{\myweb}}}
	{\scriptsize  \texttt{\href{\myblog}{\myblog}}}
\end{minipage}

\vfil

%% Name 
\noindent{\Large {\textsc{maarten steenhagen}}}
\reversemarginpar

\bigskip

%% Appointments
\medskip

\marginhead{Current Position}

\noindent\emph{University College London \vspace{0.01in}}

\ind Doctoral Candidate (PhD) in Philosophy.

\ind Dissertation research on the nature of human experience of pictorial representations. Supervisor: Prof. Mark Eli Kalderon.

\bigskip

\marginhead{AoS}

\ind Philosophy of Psychology; Philosophy of Mind \amper Language; Aesthetics.

\bigskip

\marginhead{AoC}

\ind Metaphysics; Metaethics; History of 20$^{\textrm{\tiny{th}}}$ C. Philosophy; Political Philosophy.

\bigskip
\bigskip
\vfil 
\marginhead{Education \newline (Past)}

\noindent\emph{University College London, Dept. Philosophy \vspace{0.01in}}

% \ind 2012--Present. Doctoral Candidate (PhD).

\ind 2012. MPhil Stud. research student. Dissertation: \href{http://discovery.ucl.ac.uk/1348204/}{\emph{The Collective Image}}. Supervisors: Prof. Mark Eli Kalderon; Prof. Christopher Peacocke. \vspace{-0.1in}

\bigskip

\noindent\emph{University of Utrecht, Dept. Philosophy\vspace{0.02in}}

\ind 2009. B.A., Philosophy.

\medskip

\noindent\emph{ARTeZ School of Fine Art, Dept. Monumental Art\vspace{0.02in}}

\ind 2005. BFA., Fine Art.

\bigskip

\marginhead{Appointments}

\noindent\href{http://www.heythrop.ac.uk/departments/academic-departments/philosophy/}{\emph{Heythrop College London, Dept. Philosophy} \vspace{0.01in}}

\ind 2009--Present. Tutorial Assistant in Philosophy and Psychology.

\ind 2009--Present. Research Assistant in Philosophy.

\medskip

\noindent\href{http://www.kent.ac.uk/arts/hpa/index.html}{\emph{University of Kent, History \amper Philosophy of Art} \vspace{0.01in}}

\ind 2011--2012. Assistant Lecturer, History and Philosophy of Art.

\medskip

\noindent\href{http://www.ucl.ac.uk/philosophy/}{\emph{University College London, Dept. Philosophy}} \vspace{0.01in}

\ind 2011-2012. Teaching Assistant in Philosophy.

\medskip

\noindent\href{http://www.hku.nl/web/English.htm}{\emph{Utrecht School of Arts} \vspace{0.01in}}

\ind 2008-2009. Assistant Lecturer in Art Theory and Cultural History.

\medskip

\noindent\href{http://www.uu.nl/faculty/humanities/EN/organisation/departments/departmentofphilosophy/Pages/default.aspx}{\emph{University of Utrecht, Dept. Philosophy} \vspace{0.01in}}

\ind 2007-2008. Teaching Assistant in Aesthetics and Moral Psychology.

\bigskip

% \newpage

%% Presentations
\marginhead{{\vskip 0.4em}Talks}
\medskip

\ind 2012. ``How the five die: Response to Lamb.'' University College London, 2012 Graduate Conference in Philosophy. (Response paper.)

\ind 2012. ```Non-conceptualised depiction.'' The Queen's College Oxford, Annual conference of the British Society of Aesthetics. September.

\ind 2012. ``Researching the Self.'' St. John's College Oxford, \emph{Art as a Mode of Enquiry}.

\ind 2011. ``Seeing what another would.'' Utrecht University, \emph{Expert Meeting on Philosophical Aesthetics}.

\ind 2011. ``Representing the Trinity.'' University of Ghent, Annual conference of the Dutch Society of Aesthetics.

\ind 2011. ``Truly political? A psychoanalytic understanding of engaged art.'', University of Leeds, \emph{Working Through Psychoanalysis}.

\ind 2010. ``Photography and dull causation: Response to Atencia-Linares.'' University College London, 2010 Graduate Conference in Philosophy. (Response paper.)

\ind 2010. ``Explaining the Ugly: Disharmony and Unrestrained Cognition in Kant.'' Institute of Philosophy, London-Berkeley Graduate Conference in Philosophy.

%\end{revnumerate}
\bigskip

\marginhead{{\vskip 0.3em}Publications}
\medskip

\ind  Steenhagen, M. 2010. `Explaining the Ugly,' \emph{Esthetica,} November.

\ind  Steenhagen, M. 2009. `Verwarring over lelijkheid,' \emph{Cimedart,} 39:2.

\ind  Steenhagen, M. 2008. `De keerzijde van liefde,' \emph{Cimedart,} 38:3.

\bigskip

\marginhead{{\vskip 0.4em}Grants and \newline Awards}
\medskip

\ind 2012. \href{http://www.british-aesthetics.org}{British Society of Aesthetics} PhD Studentship Award. %(£36.000)

\ind 2012. Prins Bernhard Cultuurfonds (€10.000). \emph{Not taken up.}

\ind 2012. Hendrik Müller Vaderlandsch Fonds (€5.000). \emph{Not taken up.}

\ind 2010. Dutch Association for Aesthetics Essay Contest, 2$^{\textrm{\tiny{nd}}}$ Prize.

\ind 2009. Huygens Scholarship Programme, Nuffic, The Hague (€30.000).

\bigskip

\marginhead{{\vskip 0.9em}Service to the \newline Profession}
\medskip

\ind 2012--Present. Editor of the \href{http:\\www.pjaesthetics.org}{\emph{Postgraduate Journal of Aesthetics.}}

\ind 2010--Present. Co-Organiser \href{http:\\www.londonaestheticsforum.org}{London Aesthetics Forum}, Institute of Philosophy, London.

\ind 2012--Present. Website Editor \href{http://www.philosophy-psychoanalysis.org.uk}{London Philosophy and Psychoanalysis Group}.

\ind 2012. Assistant Organiser European Society for Philosophy and Psychology (\href{http://www.eurospp.org}{ESPP}) Annual Conference 2012, Institute of Philosophy, London.

\ind 2011. UCL Representative for the \href{http://www.bppa-online.org}{British Postgraduate Philosophy Association}.

\ind 2007-2009. Editor of \href{http://www.cimedart.nl}{\emph{Cimedart}}, Philosophy Journal of the University of Amsterdam.

\ind 2008. Chair of the organisation for \emph{The Philosopher's Rally,} an international graduate conference in Philosophy, Utrecht.

\bigskip

\marginhead{{\vskip 0.9em}IT-Skills}
\medskip

\ind Microsoft Office Suite; Adobe Acrobat Professional; Adobe Photoshop and Indesign; BibDesk/Endnote and TeX-based typesetting; HTML/PHP; Open Journal Systems.

\bigskip

\marginhead{{\vskip 0.9em}Languages}
\medskip

\ind Dutch (native); English (fluent); German (good).

\rfoot{\footnotesize{\today}}

\end{document}